Kindly Bent to Free Us

### 2020-06-04 kindly/soa.tex
paragraph on Rust traits Send, Sync, and Copy

The comparison matrix in \cref{fig:comparison-matrix} gives an
overview over the systems discussed in this section. The column UAL
specifies the substructural features. Borrows and state are
self-explanatory. ``Subsumption'' indicates that unrestricted elements
can be promoted to affine and then linear. This applies to resources,
borrows, or closures. ``Polymorphism'' refers to polymorphism over
usage (unrestricted, affine, linear), notably for closures.
``Identity'' indicates that the language supports a notion
of identity, usually through existential types, as described in
\cref{identity}.
Inference is self-explanatory, formalisation
refers to the formal semantics and type soundness proof. Basis is self-explanatory.

@@

Alms~\citep{DBLP:conf/popl/TovP11} is an ML-like language with rich, kind-based
affine types and ML modules, similar to \lang.
Alms examples often rely on existential types to track the identity
of objects. For instance
\lstinline/Array.create : int -> 'a -> \E 'b. ('a, 'b) array/ where
\lstinline/'b/ uniquely identifies the array.
Due to the reliance on existentials, Alms does not support complete type inference.
Furthermore, Alms does not support borrows and often relies
on explicit capability passing.
In our experience, \affe's limited support for existential types through
regions is sufficient to express many of Alms' examples and leads to
a more convenient programming style for imperative code.
%
Alms kind structure features unions, intersections and dependent kinds while
\lang uses constrained types.
We believe most of Alms' kind signatures can be expressed equivalently in
our system: for instance the pair type constructor
has kind $\Pi\alpha\Pi\beta. \langle\alpha\rangle \sqcup \langle\beta\rangle$
(where $\alpha$ and $\beta$ are types and $\Pi$ is the dependent function)
in Alms compared to $\kvar\to\kvar\to\kvar$ in \lang thanks
to subkinding.
%
Finally, Alms provides excellent support for abstraction through
modules by allowing to keep some type unrestricted inside a module, but
exposing it as affine. \lang supports
such programming style thanks to subsumption.

@@

Rust~\citep{rust} is the first
mainstream language that builds on the concepts of borrowing and ownership
to enable safe low-level programming.
\lang is inspired by Rust's borrowing system and transfers some of its
ideas  to a functional setting with type inference, garbage collection, and
an ML-like module system.
Rust's lifetime system is more explicit and more expressive than \lang.
While Rust provides partial lifetime inference, it
does not support full type inference. 
Recently, \citet{DBLP:journals/corr/abs-1903-00982}
formalized Rust's ownership discipline, including non-lexical lifetimes.
Moreover, Rust programmers have full control over memory allocation
and memory layout of objects; they can pass arguments by value or by reference.
These features are crucial for the efficiency goals of Rust.
In contrast, \lang is garbage collected, assumes a uniform object
representation, and all arguments are passed by reference. This choice
forgoes numerous  
issues regarding interior mutability and algebraic data types.
In particular, it
allows us to easily nest mutable references inside objects, regardless
whether they are linear or unrestricted.

>>>>

The comparison matrix in \cref{fig:comparison-matrix} gives an
overview over the systems discussed in this section. The column UAL
specifies the substructural features supported (unrestricted, affine,
linear). Borrows and state indicate support for the respective
feature. ``Subsumption'' indicates that unrestricted elements 
can be promoted to affine and then linear. This promotion applies to
all objects, resources,
borrows, or closures. ``Polymorphism'' refers to polymorphism over
substructural features, notably for closures.
``Identity'' indicates that the language supports a notion
of identity, usually through existential types, as described in
\cref{identity}.
Inference is self-explanatory, formalization
refers to the existence of a formal semantics and type soundness proof. Basis is self-explanatory.

@@

Alms~\citep{DBLP:conf/popl/TovP11} is an ML-like language with rich, kind-based
affine types and ML modules, similar to \lang.
Alms examples often rely on existential types to track the identity
of objects. For instance, consider the signature
\lstinline/Array.create : int -> 'a -> \E 'b. ('a, 'b) array/ where
\lstinline/'b/ uniquely identifies the array.
Due to the reliance on existentials, Alms does not support complete type inference.
Furthermore, Alms does not support borrows and often relies
on explicit capability passing.
In our experience, \affe's limited support for existential types through
regions is sufficient to express many of Alms' examples and leads to
a more convenient programming style for imperative code.
%
Alms kind structure features unions, intersections and dependent kinds while
\lang uses constrained types.
We believe most of Alms' kind signatures can be expressed equivalently in
our system: for instance the pair type constructor
has kind $\Pi\alpha\Pi\beta. \langle\alpha\rangle \sqcup \langle\beta\rangle$
(where $\alpha$ and $\beta$ are types and $\Pi$ is the dependent function)
in Alms compared to $\kvar\to\kvar\to\kvar$ in \lang thanks
to subkinding.
%
Finally, Alms provides excellent support for abstraction through
modules by allowing to keep some type unrestricted inside a module, but
exposing it as affine. \lang supports
such programming style thanks to subsumption.

@@

Rust~\citep{rust} is the first
mainstream language that builds on the concepts of borrowing and ownership
to enable safe low-level programming.
\lang is inspired by Rust's borrowing system and transfers some of its
ideas  to a functional setting with type inference, garbage collection, and
an ML-like module system.
Instead of kinds, Rust uses marker traits like \lstinline/Copy/,
\lstinline/Send/, and \lstinline/Sync/ to distinguish characteristics 
of types, which would just be unrestricted in \lang.
Also, Rust's lifetime system is more explicit and more expressive than \lang.
While Rust provides partial lifetime inference, it
does not support full type inference. 
Recently, \citet{DBLP:journals/corr/abs-1903-00982}
formalized Rust's ownership discipline, including non-lexical lifetimes.
Moreover, Rust programmers have full control over memory allocation
and memory layout of objects; they can pass arguments by value or by reference.
These features are crucial for the efficiency goals of Rust.
In contrast, \lang is garbage collected, assumes a uniform object
representation, and all arguments are passed by reference. This choice
forgoes numerous  
issues regarding interior mutability and algebraic data types.
In particular, it
allows us to easily nest mutable references inside objects, regardless
whether they are linear or unrestricted.
